### 2019-09-16 presentations/2018-05-25_ica/presentation.Rmd

---
title: "Longitudinal Analysis of the Privacy Paradox"
author: "Tobias Dienlin (@tdienlin), Philipp Masur (@masurphil), Sabine Trepte <br /> <br /> University of Hohenheim"
date: "2018-05-25"
output:
  xaringan::moon_reader:
    css: "theme.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
    self_contained: false
---

layout: true

<div class="my-footer"></div>

```{r load_libraries, echo=FALSE, cache=FALSE, include=FALSE}
pks <- c("RefManageR", "tidyverse")
invisible(lapply(pks, library, character.only = TRUE))
load("../../.RData")
```

```{r load_refs, echo=FALSE, cache=FALSE}
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = "year", 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
my_bib <- ReadBib("../../bibliography/references.bib", check = FALSE)
```

---

# Brief history of privacy paradox

- Privacy concerns **inconsistent with** SNS behavior `r Cite(my_bib, "Acquisti.2003")`

--

- Description of SNS privacy behavior as **paradoxical** `r Cite(my_bib, "Barnes.2006")`

--

- Afterward, explicit studies on privacy paradox

   - Some show **no relation** between privacy concerns and information sharing  `r Cite(my_bib, "Taddicken.2014")`

   - Some show **significant relation** between privacy concerns and information sharing `r Cite(my_bib, "Dienlin.2016a")`

--

- Meta-Analysis: Privacy concerns and information sharing on SNSs are **significantly related, _r_ = -.13** `r Cite(my_bib, "Baruh.2017")`

---

# Aims of current study

- **Longitudinal analysis** of privacy concerns and information sharing

--

- Separation of **between and within-person** variance

--

    - General differences between people?

--

    - Short term relations within a person?

--

    - Long term relations within a person?

--

- Longitudinal analysis of **privacy attitudes**

---

background-image: url(../../figures/model_1.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/model_2.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/model_3.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/model_4.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/model_5.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/model_7.png)
background-size: 850px
background-position: 50% 10%

---

class: center, middle, inverse

# Method

---

# Procedure & participants

- Paper-pencil questionnaire

    - **3 waves**: May 2014, November 2014, May 2015

- **Representative sample** of the German population

    - _N_ = 1,419

    - Age = 54 years, _SD_ = 15 years

    - Female = 49%

---

# Measures

.smaller[
**Privacy concerns** (6 items; CFI = .97, RMSEA = .06, $\omega$ = .90)

- “Are you concerned that institutions or secret services could collect and analyze the data that you are disclosing in the Internet?”
- “Are you concerned about people you do not know obtaining personal information about you from your online activities?”

**Attitudes information sharing** (10 items; CFI = .96, RMSEA = .11, $\omega$ = .94)

- „Do you think that it is sensible to share the following pieces of information online?”
- 1. Second name, 2. First name, 3. Residence (town), 4. Street + no., 5. Financial information, 6. Medical information, 7. e-mail address, 8. Phone number, 9. Job, 10. Education

**Information sharing** (10 items; CFI = .97, RMSEA = .10, $\omega$ = .95)

- How often do you share the following pieces of information online?
- 1. Second name, 2. First name, 3. Residence (town), 4. Street + no., 5. Financial information, 6. Medical information, 7. e-mail address, 8. Phone number, 9. Job, 10. Education
]

---

# Statistics

- Hypotheses tested with **one-tailed** approach, Research Questions with **two-tailed** approach (Holm-Bonferroni-corrected)

--

- Smallest effect size of interest `r Cite(my_bib, "Lakens.2014")`: $\beta$ = **.10**

--

- Alpha and beta-errors **balanced** `r Cite(my_bib, "Rouder.2016")`

--

   - $\alpha$ = `r alpha_balanced %>% round(3)`
   - $\beta$ = `r alpha_balanced %>% round(3)`
   - Power = `r (1 - alpha_balanced) * 100 %>% round(3)`%

--

- Data analysed with Random Intercept-Cross Lagged Panel Model (RI-CLPM) `r Cite(my_bib, "Hamaker.2015")`

---

class: center, middle, inverse

# Results

---

background-image: url(../../figures/results_1.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/results_2.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/results_3.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/results_4.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/results_6.png)
background-size: 850px
background-position: 50% 10%

---

background-image: url(../../figures/results_7.png)
background-size: 850px
background-position: 50% 10%

---

class: center, middle, inverse

# Discussion

---

# Implications

- Important to conduct longitudinal studies to separate between- and within relations

--

- Between people

    - Concerns and sharing **related weakly**
    - Attitudes and sharing **related strongly**

--

- Within people

    - Concerns and sharing **related weakly**
    - Attitudes and sharing **related substantially**
    - **No long-term effects** for concerns and sharing

--

- Future research: Shorter invervals with actual observations of behavior

--

- Information sharing online **not paradoxical**

    - Instead of privacy paradox, rather **privacy orthodox**

---

# References

.small[
```{r, 'refs', results='asis', echo=FALSE}
PrintBibliography(my_bib)
```
]

---

class: center, middle
background-image: url(../../figures/bmbf.png)
background-position: center bottom 50px
background-size: 30%

# Thank you for your attention

This research was funded by the German Federal Ministry of Education and Research (BMBF) Grant 16KIS0094