### 2020-04-23 index.Rmd
setting up github website

---
title: "Longitudinal Analysis of the Privacy Calculus"
---

On this website, you can find the companion material for the paper _A Longitudinal Analysis of the Privacy Paradox_.

You can either read the paper [online](manuscript.html) or download it as [pdf](manuscript.pdf). You can also find the detailed code and summaries of the [analyses](analyses.html) that were reported in the paper.

Please note that this OSM does not contain the actual data, which can be downloaded from http://dx.doi.org/10.7802/1937.

This website is produced directly from the project's [github repository](https://github.com/tdienlin/privacy_calc_exp). You can download the material, file an issue, or submit a pull request.