### 2021-03-16 items.Rmd

---
title: "List of Items"
output: 
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
---

# Items
## Privacy Concern

The following deals with things that can happen on the Internet.

 - Are you concerned about people you do not know obtaining personal information about you from your online activities?
 - Are you concerned that if you use your credit card to buy something on the Internet your credit card number will be obtained/intercepted by someone else?
 - Are you concerned that an email you sent may be read by someone else besides the person you sent it to?
 - Are you concerned that institutions or secret services could collect and analyze the data that you are disclosing on the Internet?
 - Are you concerned that website providers are tracking your online activities?
 - Are you concerned that you don’t know what other organizations or website providers are doing with you data?

(NA = "no Internet user / no reply", 1 = "Not concerned at all" : 5 = "very concerned")

## Attitude Toward Online Sharing of Personal Information

Now we will deal with information sharing on the Internet. Do you think that it is sensible to disclose the following pieces of information online, i.e., via the Internet?

 - Financial information (bank data, account balances…)
 - Medical information (health, medical records…)
 - Second name
 - First name
 - Residence
 - Street and house number
 - Email address
 - Phone number
 - Job (past and present)
 - Education

(NA = “no Internet user / no reply”, 1 = "not sensible at all", 2 = "somewhat not sensible", 3 = "in part", 4 = "somewhat sensible", 5 = "very sensible")

## Online Sharing of Personal Information

In the following, please indicate how often you disclose the following information on the Internet. How often do you disclose the following pieces of information online, i.e., on the Internet?

 -  Financial information (bank data, account balances…)
 -  Medical information (health, medical records…)
 -  Second name
 -  First name
 -  Residence
 -  Street and house number
 -  Email address
 -  Phone number
 -  Job (past and present)
 -  Education

(NA = “no Internet user / no reply”, 1 = "every day", 2 = "weekly", 3 = "monthly", 4 = "less than that", 5 = "never")